---
title: "Squash Miranda's Misery"
output: html_notebook
---


Miranda needs a correct survival plot for germination time but R is punishing her. 

Time to help!


```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

```{r}
boo <- read_csv("cox_germ_soil8-2-21.csv")
```

What we gotta plot: 

x = time to germination (column = DaysToGerm)
y = germination probability (column = PlotPropGerm)
color = strata (column = Habitat)


# Germination Rate

If you want to change Forest to Natural: 
 %>% 
 mutate(Treatment = case_when(
    Treatment == "Forest" ~ paste("Natural"),
    Treatment != "Forest" ~ paste(Treatment)
  ))


## Habitat and Treatment

1. DaystoGerm should be 0 if actually 0, Censor should be NA for DaystoGerm at 0 - ended up filtering off this line


2. now lets select the columns we want for plotting so we can make sure there is only one value per time to germination and habitat

```{r}

no_days <- boo %>% 
  filter(DaysToGerm != 0) %>% 
  mutate(plot_treat = paste(Habitat, Treatment, sep="/")) %>% 
  group_by(plot_treat) %>% 
  count(DaysToGerm) %>% 
  mutate(cums = cumsum(n)) %>% 
  mutate(prop = cums/400) 
  

germd_gg <- ggplot(no_days)+
  geom_line(aes(x=DaysToGerm, y=prop, color = plot_treat), size = 1)+
  theme_classic()+
  ylim(0,1)+
  scale_x_continuous(name = "Days to Germination", limits = c(0,12), breaks = seq(0,12,1))+
  labs(y="Proportion Germinated", color = "Habitat/Treatment")
germd_gg

ggsave(plot = germd_gg, file = "../figures/propgermd.png", width = 20, height = 25, units = "cm", dpi = 800)
```


## Population

```{r}
no_days1 <- boo %>% 
  filter(DaysToGerm != 0) %>% 
  mutate(plot_treat = paste(Population, Treatment, sep="/")) %>% 
  group_by(plot_treat) %>% 
  count(DaysToGerm) %>% 
  mutate(cums = cumsum(n)) %>% 
  mutate(prop = cums/50) 
  

germd1_gg <- ggplot(no_days1)+
  geom_line(aes(x=DaysToGerm, y=prop, color = plot_treat), size = 1)+
  geom_point(aes(x=DaysToGerm, y=prop, color = plot_treat, shape = plot_treat), size = 5)+
  theme_classic()+
  ylim(0,1)+
  scale_x_continuous(name = "Days to Germination", limits = c(0,12), breaks = seq(0,12,1))+
  scale_shape_manual(values = c(16, 17, 1, 2, 16, 17, 1, 2,
                                16, 17, 1, 2, 16, 17, 1, 2,
                                16, 17, 1, 2, 16, 17, 1, 2,
                                16, 17, 1, 2, 16, 17, 1, 2))+
  labs(y="Proportion Germinated", color = "Population/Treatment", shape = "Population/Treatment")
germd1_gg

ggsave(plot = germd1_gg, file = "../figures/propgermd1.png", width = 20, height = 25, units = "cm", dpi = 800)
```


# Germination Success


## Habitat and Treatment

```{r}

suc <- boo %>% 
  filter(DaysToGerm != 0) %>% 
  mutate(plot_treat = paste(Habitat, Treatment, sep="/")) %>% 
  group_by(Habitat, Treatment) %>% 
  count(DaysToGerm) %>% 
  mutate(cums = cumsum(n)) %>% 
  mutate(prop = cums/400) 

suc_gg <- ggplot(suc)+
  geom_boxplot(aes(x=Treatment, y=prop, fill = Habitat), size = 0.5)+
  theme_classic()+
  ylim(0,1)+
  scale_fill_manual(values = c("forestgreen","gray58"))+
  labs(y="Proportion Germinated", fill = "Habitat", x="Treatment")
suc_gg

ggsave(plot = suc_gg, file = "../figures/propgerms.png", width = 20, height = 25, units = "cm", dpi = 800)

sucf_gg <- ggplot(suc)+
  geom_boxplot(aes(x=Treatment, y=prop, fill = Habitat), size = .5)+
  theme_classic()+
  ylim(0,1)+
  facet_grid(Habitat~.)+
  scale_fill_manual(values = c("forestgreen","gray58"))+
  labs(y="Proportion Germinated", fill = "Habitat", x="Treatment")
sucf_gg

ggsave(plot = sucf_gg, file = "../figures/propgermsF.png", width = 20, height = 25, units = "cm", dpi = 800)
```

## Population and Treatment

```{r}
suc1 <- boo %>% 
  filter(DaysToGerm != 0) %>% 
  mutate(plot_treat = paste(Population, Treatment, sep="/")) %>% 
  group_by(Population, Treatment, Habitat) %>% 
  count(DaysToGerm) %>% 
  mutate(cums = cumsum(n)) %>% 
  mutate(prop = cums/50) 

suc1_gg <- ggplot(suc1)+
  geom_boxplot(aes(x=Treatment, y=prop, fill = Population), size = 0.5)+
  theme_classic()+
  ylim(0,1)+
  facet_grid(Habitat~.)+
  #scale_fill_manual(values = c("forestgreen","gray58"))+
  labs(y="Proportion Germinated", fill = "Population", x="Treatment")
suc1_gg

ggsave(plot = suc1_gg, file = "../figures/propgerms1.png", width = 20, height = 25, units = "cm", dpi = 800)
```


# Response to Soil Type and Competition


## Reformatting datasheet

```{r}
compt <- read_csv("../competition_biomass.csv")

# change "Empty" Competition entries to "None"
compta <- compt %>% 
  mutate(Competition = case_when(
    Competition == "Empty" ~ paste("None"),
    Competition != "Empty" ~ paste(Competition)
  )) %>% 
  mutate(Habitat = case_when(
    Habitat == "Off-Roadsideoads" ~ paste("Off-road"),
    Habitat != "Off-Roadsideoads" ~ paste(Habitat)
  ))  %>% 
  replace_na(list(Dittrichia_biomass = 0, Grass_biomass = 0)) %>% 
  mutate(Dittrichia_biomass1000 = Dittrichia_biomass*10000, Grass_biomass1000 = Grass_biomass*10000) %>% 
  filter(Dittrichia_biomass != 0) %>% 
  mutate(log_dic = log(Dittrichia_biomass1000)) %>% 
  mutate(dic_mg = Dittrichia_biomass*100)
  

```


## Habitat/Treatment Plant Biomass

```{r}

compta_gg <- ggplot(compta)+
  geom_boxplot(aes(x=Competition, y=log_dic, fill = Soil))+
  theme_classic()+
  scale_fill_manual(values = c("gray58", "forestgreen"))+
  labs(y="log(Biomass)", fill = "Soil", x="Competition")
compta_gg

ggsave(plot = compta_gg, file = "../figures/compt_biomass.png", width = 20, height = 25, units = "cm", dpi = 800)
```

## Populations/Soil + Habitat

```{r}
compta1_gg <- ggplot(compta)+
  geom_boxplot(aes(x=Soil, y=log_dic, fill = Population))+
  theme_classic()+
  facet_grid(Habitat~.)+
  #scale_fill_manual(values = c("gray58", "forestgreen"))+
  labs(y="log(Biomass)", fill = "Population", x="Soil")
compta1_gg

ggsave(plot = compta1_gg, file = "../figures/compt1_biomass.png", width = 20, height = 25, units = "cm", dpi = 800)
```


## Biomasses Scatterplot


```{r}
scatter <- compta %>% 
  filter(Competition != "None") %>%
  filter(Grass_biomass != 0) %>% 
  mutate(scx = log(Grass_biomass), scy = log(Dittrichia_biomass)) %>% 
  mutate(type = paste(Soil, Competition, sep="/"))

scbio_gg <- ggplot(scatter)+
  geom_point(aes(x=scx, y=scy, shape = Soil, color = Competition))+
  scale_color_manual(values = c("orange", "purple"))+
  geom_smooth(method=lm, aes(x=scx, y=scy, linetype = Soil), color = "black")+
  theme_classic()+
  labs(y = "Dittrichia biomass (g)", x ="Grass biomass (g)", shape = "Soil", color = "Competition")

scbio_gg

ggsave(plot = scbio_gg, file = "../figures/scat_biomass.png", width = 20, height = 25, units = "cm", dpi = 800)
```





