---
title: "Relative fitness in a field setting"
output: html_notebook
date: December 2022
authors: Miranda Melen, Nicky Lustenhouwer
---
#Background
At Blue Oak Ranch Reserve, we established a 10m x 26m fenced field site to test whether evolution over the course of invasion away from roads has resulted in enhanced performance in undisturbed vegetation relative to roadside populations. The experiment was replicated in a randomized block design (20 plots in total). Each plot was 1.5m2 with 16 D. graveolens growing in a 4 x 4 grid centered on the plot. There was 33cm between each plant and 25cm between the edge plants and the border of the plot.

The experiment included multiple treatments; however, only the two most relevant to the focus of this paper are included here. We tested whether plant genotypes collected from the two habitats responded differently to the disturbance of biomass removal. We tilled in December 2020 to completely remove below and aboveground biomass, and then weeded to remove aboveground biomass throughout the growing season. In contrast, we left the control plots untouched, allowing the previous year’s thatch to persist and background vegetation to grow throughout the experiment. 

In January 2021, we germinated seeds in Petri dishes at the UCSC Coastal Science Campus greenhouse incubation chambers before transplanting them into field-collected soil (collected in late December 2020 from Blue Oak Ranch Reserve). Seedlings grew in the greenhouse for about eight weeks until all plants had their first two true leaves emerge and lengthen. Ideally, we would have placed seeds directly into the field, but to maximize biosafety, we used seedling transplants that could be tracked with 100% certainty.

We measured the longest leaf for each plant and then transplanted them into the ground in late February 2021 at Blue Oak Ranch Reserve. During the first month of growth, we replaced any D. graveolens that died. We conducted weekly phenology surveys to assess D. graveolens plant health, and at the first sign of buds, we measured plant height and harvested the aboveground biomass by cutting at the root crown and drying in a 60ºC oven for 3 days before weighing.

#Data Analysis
Statistical analyses were performed in R version 4.2.1 (R Core Team 2022) using linear mixed-effects models with the lme4 (Bates et al. 2015), lmerTest (Kuznetsova et al. 2017), and DHARMa packages (Hartig 2022), generalized linear mixed models with the glmmTMB package (Brooks et al. 2017), and mixed effects cox models with the coxme (Therneau 2022a) and survival (Therneau 2022b) packages.

##Models Included

###Cox proportional hazard models
Here we will use 'coxme' which allows you to conduct mixed effects Cox proportional hazards models. Information here: https://cran.r-project.org/web/packages/coxme/vignettes/coxme.pdf. We conducted a germination experiment using Dittrichia graveolens seeds on filter paper. "Surv" creates a survival object to combine the days column (NumDaysAlive) and the censor column (Censor) to be used as a response variable in a model formula.The model follows the same syntax as linear models (lm, lmer, etc). Fixed effects: "var1 * var2" will give you the interaction term and the individual variables: so "var1 + var2 + var1:var2". To add random effects, type "+ (1|random effect)".

Assumptions for cox models: https://www.theanalysisfactor.com/assumptions-cox-regression/

###Linear mixed-effects models
Linear Mixed Effects models are used for regression analyses involving dependent data. For a tutorial: https://doi.org/10.1177/2515245920960351

###Generalized linear mixed models



#Libraries
```{r}
#install.packages("coxme")
#install.packages("survival")
#install.packages("ggplot2")
#install.packages("ggfortify")
#install.packages("car")
#install.packages("multcomp")
#install.packages("lme4")
#install.packages("lmerTest")
#install.packages("DHARMa")
#install.packages("dplyr")
#install.packages("emmeans")
#install.packages('TMB', type = 'source')
#install.packages("glmmTMB")
#install.packages("MASS")
#install.packages("emmeans")
#install.packages("AICcmodavg")
library(coxme)
library(survival)
library(ggplot2)
library(ggfortify)
library(car)
library(multcomp)
library(lme4)
library(lmerTest)
library(DHARMa)
library(dplyr)
library(emmeans)
library(TMB)
library(glmmTMB)
library(MASS)
library(emmeans)
library(AICcmodavg)
```

#Load Data
This dataframe has one row per plant (800 observations). Data are for survivorship curves (3 censor options), the number of days the plant stayed alive (NumDaysAlive) and aboveground biomass. Censors with a 1 denote reaching the event (CensorAll = died, CensorBiomass = survived to collect biomass, CensorReproduction = survived to reproduce) and a 0 denoting when a seed didn't germinate by the last census date (Census = 11/15/21). CensorReproduction will be most useful in understanding the amount of biomass produced by an individual when buds appear.
```{r}
mydata<-read.csv("/Users/Miranda/Documents/Education/UC Santa Cruz/Dittrichia/Dittrichia_Analysis/data/field_relative_fitness/field_relative_fitness_2021.csv",stringsAsFactors=T)
str(mydata) #Check that each column has the right class (factor, integer, numeric, etc.)
mydata$Site<-as.character(mydata$Site)
mydata$Treatment<-factor(mydata$Treatment,levels=c("Control","Biomass Removal","Hemizonia","Raking","Raking + Clipping")) #Changing the contrast order so that everything is compared to Control
```

#Early Growth
This code uses Growth data with Habitat (roadside and off-road) and Treatment in a glmm model. Anova and Tukey tests are used on the successful Model4 with the creation of a box plot as a finished product.

Note: 10 blocks, 5 treatments, 16 populations (CHE-O), 8 sites (random: population pairs; CHE), 2 habitat (fixed effect: roadside and off-road; R or O)

Also Note: Growth data is a measure of growth of longest leaf. Each plant was measured upon transplanting into the field in March and then again in June 2021. This plant has a juvenile stage of a basal rosette, and then it bolts and produced smaller cauline leaves. The goal was to capture early growth data for this plant before bolting occurs so that the measurements capture the basal rosette stage, however in some cases the plants bolted earlier than expected and the resulting measurement was smaller than previous. In these cases we determined that the data should be removed from the dataset as the negative number (or changing it to a zero) does not reflect the biological importance of the measurement.

Here we need to filter the data to remove Growth>0 and to convert plant measurement dates to date format
```{r}
growth_mydata<-mydata%>%filter(Growth>0) #Here I am only looking at the Growth data that is greater than 0 (see Also Note above)
growth_mydata$PlantDate<-as.Date(growth_mydata$PlantDate,"%m/%d/%y") 
growth_mydata$Num.Days.Growth<-as.Date("2021-05-22")-growth_mydata$PlantDate
growth_mydata$Num.Days.Growth<-as.numeric(growth_mydata$Num.Days.Growth)
growth_mydata$Growth.Rate<-growth_mydata$Growth/growth_mydata$Num.Days.Growth #First calculate number of days of growth to get the Rate (Growth/Num.Days.Growing). Then fit data to Beta distribution
```

##Histograms
Original data
```{r}
hist(growth_mydata$Growth.Rate,col='steelblue',main='Original') #Original data is skewed, let's test for normality and consider log transforming the data
shapiro.test(growth_mydata$Growth.Rate)
```

Log transform data (https://www.statology.org/transform-data-in-r/)
```{r}
log_growth_mydata<-log10(growth_mydata$Growth.Rate)
hist(log_growth_mydata,col='coral2',main='Log Transformed') #Log transformed data, this looks better than the original distribution
shapiro.test(log_growth_mydata) #Data does not improve with log transformation
```

Next I tried a square root transformation, which improved the distribution, but my first model failed the singular fit (see Model 1).
```{r}
sqrt_growth_mydata<-sqrt(growth_mydata$Growth.Rate)
hist(sqrt_growth_mydata,col='coral2',main='Log Transformed') #Square root transformed data looks better than the original distribution
shapiro.test(sqrt_growth_mydata)
```

##Model 1
```{r}
growth_fullmodel1<-lmer(sqrt(Growth.Rate)~Habitat*Treatment+(1|Site)+(1|Block),data=growth_mydata)
isSingular(growth_fullmodel1,tol=1e-4) #=True
summary(growth_fullmodel1) #Variance explained by Site = 0.000
anova(growth_fullmodel1)
```

Site as a random effect does not explain any of the variance in the model, therefore let's try Site as a fixed effect to see if it adds to the model.

##Model 2 - Modeling Site as a fixed effect
```{r}
growth_fullmodel2<-lmer(log(Growth.Rate)~Habitat*Treatment+Site+(1|Block),data=growth_mydata)
summary(growth_fullmodel2) #As a fixed effect, one of the Sites (OAP) is significant.
anova(growth_fullmodel2) #Site as a fixed effect accounts for 3% of the variance
```

Now we'll look at the QQ plots and the residuals using DHARMa
```{r}
qqnorm(resid(growth_fullmodel2)) #qqplot
qqline(resid(growth_fullmodel2)) #add the line
testDispersion(growth_fullmodel2) #red line should be in the middle of the distribution
myDHARMagraph2<-simulateResiduals(growth_fullmodel2) #making a graph using DHARMa package, also testing for heteroscedasticity, https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html#heteroscedasticity
plot(myDHARMagraph2) #plotting graph. At this point, you don't want any text or lines to be red. The QQ plots are not looking good for this model. Let's try fitting to a negative binomial distribution.
```

##Model 3 - Negative Binomial
```{r}
#growth_fullmodel3<-glmer.nb(Growth.Rate~Habitat*Treatment+(1|Site)+(1|Block),data=growth_mydata)
summary(growth_fullmodel3)
qqnorm(resid(growth_fullmodel3)) #qqplot
qqline(resid(growth_fullmodel3)) #add the line
testDispersion(growth_fullmodel3) #red line should be in the middle of the distribution
myDHARMagraph3<-simulateResiduals(growth_fullmodel3) #making a graph using DHARMa package, also testing for heteroscedasticity, https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html#heteroscedasticity
plot(myDHARMagraph3) #plotting graph.The QQ plots are not looking good for this model.
```



So this model isn't working well either. Let's try building a model and fitting it to a Beta distribution, first without a link function. Check it with DHARMa, and if it doesn't look good, then fit it to a Beta distribution with a logit link function.

Model 4 - Beta distribution####
#Now I'm using glmm because I'm fitting to other distributions
fullmodel4<-glmmTMB(Rate~Habitat*Treatment+(1|Site)+(1|Block),family=beta_family(),data=mydata)
summary(fullmodel4)
qqnorm(resid(fullmodel4)) #qqplot
qqline(resid(fullmodel4)) #add the line
testDispersion(fullmodel4) #red line should be in the middle of the distribution
myDHARMagraph4<-simulateResiduals(fullmodel4) #making a graph using DHARMa package, also testing for heteroscedasticity, https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html#heteroscedasticity
plot(myDHARMagraph4) #plotting graph. Looks good, but check the outliers to make sure they are real.

Checking outliers:
ggplot(data=mydata,aes(x=Habitat,y=Rate))+geom_boxplot()
max(mydata$Rate[mydata$Habitat== "Off-road"])
max(mydata$Rate[mydata$Habitat== "Roadside"]) 
Yes, these outliers make sense, so I don't need to worry about the red stars in the DHARMa plot.

Here is the best model
summary(fullmodel4)

Post-Hoc Test####
Remove non-significant interaction terms before running the Tukey.

?emmeans, emmeans(model, pairwise ~ treatment)
fullmodel4.1<-glmmTMB(Rate~Treatment+(1|Site)+(1|Block),family=beta_family(),data=mydata)
summary(fullmodel4.1)
emmeans(fullmodel4.1,pairwise~Treatment)

Plotting Successful Model###
ggplot(data=mydata,aes(x=Treatment,y=Rate))+geom_boxplot() #plot data from log(data)
ggplot(data=mydata,aes(x=Habitat,y=Rate))+geom_boxplot() #plot data from log(data)

##Historgrams
##Models
###Best Model
##ggplot



#Survival Analysis
##Histograms
```{r}
#Number of Days Alive - All data
hist(mydata$NumDaysAlive,col='steelblue',main='Original') 

#Number of Days Alive - By Habitat
ggplot(mydata,aes(x=NumDaysAlive))+geom_histogram()+facet_wrap(vars(Habitat)) #Here we see that both Habitats have the same bi-modal distribution.

#Number of Days Alive - By Treatment
ggplot(mydata,aes(x=NumDaysAlive))+geom_histogram()+facet_wrap(vars(Treatment)) #Here we see that 4 of the treatments have the same bi-modal distribution and Biomass Removal is right skewed.
```

##Cox Model
Start by making a simple model with no random effects. This will be compared to the full model with random effects.
```{r}
cox_simplemodel1<-coxph(Surv(NumDaysAlive,CensorReproduction)~Habitat*Treatment,data=mydata)
summary(cox_simplemodel1)
print(cox_simplemodel1)
predict(cox_simplemodel1)
hist(predict(cox_simplemodel1))
```

Now make a full model using random effects
```{r}
cox_fullmodel1<-coxme(Surv(NumDaysAlive,CensorReproduction)~Habitat*Treatment+(1|Site)+(1|Block),data=mydata)
summary(cox_fullmodel1)
print(cox_fullmodel1)
predict(cox_fullmodel1)
hist(predict(cox_fullmodel1))
```

Now we can compare the models to see which model is best
```{r}
anova(cox_simplemodel1,cox_fullmodel1) 
#See example: https://www.rdocumentation.org/packages/coxme/versions/2.2-16/topics/coxme
AIC(cox_simplemodel1,cox_fullmodel1) #But comparing the AIC scores is easiest. Keep the lower AIC score because that is considered the better model. Here it is the fullmodel1.
```

Now, let's make a model with no interaction term
```{r}
cox_fullmodel2<-coxme(Surv(NumDaysAlive,CensorReproduction)~Habitat+Treatment+(1|Site)+(1|Block),data=mydata) 
summary(cox_fullmodel2)
```

Let's compare the the first two models to test for the significance of the term that is removed (using LR)
```{r}
anova(cox_fullmodel1,cox_fullmodel2) #Not significant
AIC(cox_fullmodel1,cox_fullmodel2) #Interaction term is not significant and the second model has a lower AIC score. So we can drop the interaction term and keep fullmodel2.
```

So, now let's add in population nested under site as a random effect
```{r}
cox_fullmodel3<-coxme(Surv(NumDaysAlive,CensorReproduction)~Habitat+Treatment+(1|Site/Population)+(1|Block),data=mydata) 
summary(cox_fullmodel3)
```

Now we can compare the models to see which model is best
```{r}
anova(cox_fullmodel2,cox_fullmodel3) #Significant
AIC(cox_fullmodel2,cox_fullmodel3) #Looks like fullmodel3 is the better model because of the lower AIC score
```

###Best Model
```{r}
#cox_fullmodel3<-coxme(Surv(NumDaysAlive,CensorReproduction)~Habitat+Treatment+(1|Site/Population)+(1|Block),data=mydata) 
summary(cox_fullmodel3)
```

###Risk Assessment
These values are found in the model summary, but if you want to pull them out, here is how you interpret them.
```{r}
#1 = no effect, <1 = decreased risk of death, >1 = increased risk of death.

exp(coef(cox_fullmodel3)) #This should be interpreted that Biomass Removal is almost 5% more likely to survive to reproduction compared to Control, and Raking + Clipping is about 2.5% more likely to survive to reproduction compared to Control.
exp(ranef(cox_fullmodel3)$Block)
exp(ranef(cox_fullmodel3)$Site) # Pretty even among all the sites
Anova(cox_fullmodel3)
```

Looks like roadside and offroad plants are the same, so let's combine them together in a model (aka, removing the Habitat term)
```{r}
cox_fullmodel4<-coxme(Surv(NumDaysAlive,CensorReproduction)~Treatment+(1|Site/Population)+(1|Block),data=mydata) 
summary(cox_fullmodel4)
summary(cox_fullmodel3)
anova(cox_fullmodel3,cox_fullmodel4) #Significant
AIC(cox_fullmodel3,cox_fullmodel4) #Looks like fullmodel4 is the better model because the AIC score is within 2 points of each other, therefore the models are assessed the same and you should take the simpler model. But this is for another manuscript, probably.
```
##ggplot - Survival Curves
Plot by habitat
```{r}
cox_model_graph1<-survfit(Surv(NumDaysAlive,CensorReproduction)~Habitat,data=mydata)
autoplot(cox_model_graph1)+labs(x="\n Survival Time (Days)",y="Survival Probabilities\n",title="Survival Times Of \n Roadside and Off-road Populations\n")+theme(plot.title=element_text(hjust=0.5),axis.title.x=element_text(face="bold",color="Black",size = 12),axis.title.y=element_text(face="bold",color="Black",size=12),legend.title=element_text(face="bold",size=10))
```

Plot by all treatment
```{r}
cox_model_graph2<-survfit(Surv(NumDaysAlive,CensorReproduction)~Treatment,data=mydata)
autoplot(cox_model_graph2)+labs(x="\n Survival Time (Days)",y="Survival Probabilities\n",title="Survival Times Of \n Roadside and Off-road Populations\n")+theme(plot.title=element_text(hjust=0.5),axis.title.x=element_text(face="bold",color="Black",size = 12),axis.title.y=element_text(face="bold",color="Black",size=12),legend.title=element_text(face="bold",size=10))
```
Post hoc Tukey
```{r}
summary(glht(cox_fullmodel4, mcp(Treatment="Tukey")))
```

Plot by 2 treatments (Control and Biomass Removal), and adding in Habitat
```{r}
cox_sub_mydata<-subset(mydata,Treatment%in%c('Control','Biomass Removal'))
cox_model_graph3<-survfit(Surv(NumDaysAlive,CensorReproduction)~Treatment,data=sub_mydata)
autoplot(cox_model_graph3)+labs(x="\n Survival Time (Days)",y="Survival Probabilities\n",title="Survival Times Of \n Roadside and Off-road Populations\n")+theme(plot.title=element_text(hjust=0.5),axis.title.x=element_text(face="bold",color="Black",size = 12),axis.title.y=element_text(face="bold",color="Black",size=12),legend.title=element_text(face="bold",size=10))

cox_model_graph3<-survfit(Surv(NumDaysAlive,CensorReproduction)~Treatment+Habitat,data=sub_mydata)
autoplot(cox_model_graph3)+labs(x="\n Survival Time (Days)",y="Survival Probabilities\n",title="Survival Times Of \n Roadside and Off-road Populations\n")+theme(plot.title=element_text(hjust=0.5),axis.title.x=element_text(face="bold",color="Black",size = 12),axis.title.y=element_text(face="bold",color="Black",size=12),legend.title=element_text(face="bold",size=10))
```