---
title: "Relative fitness in a field setting - Figure 4 (Evolution MS)"
output: html_notebook
date: December 2022
author: Miranda Melen
contributors: Laura Goetz, Matt Kustra, Nicky Lustenhouwer
---
#Background
At Blue Oak Ranch Reserve, we established a 10m x 26m fenced field site to test whether evolution over the course of invasion away from roads has resulted in enhanced performance in undisturbed vegetation relative to roadside populations. The experiment was replicated in a randomized block design (20 plots in total). Each plot was 1.5m2 with 16 D. graveolens growing in a 4 x 4 grid centered on the plot. There was 33cm between each plant and 25cm between the edge plants and the border of the plot.

The experiment included multiple treatments; however, only the two most relevant to the focus of this paper are included here. We tested whether plant genotypes collected from the two habitats (roadside and vegetated) responded differently to the disturbance of biomass removal. We tilled in December 2020 to completely remove below and aboveground biomass, and then weeded to remove aboveground biomass throughout the growing season. In contrast, we left the control plots untouched, allowing the previous year’s thatch to persist and background vegetation to grow throughout the experiment. 

In January 2021, we germinated seeds in Petri dishes at the UCSC Coastal Science Campus greenhouse incubation chambers before transplanting them into field-collected soil (collected in late December 2020 from Blue Oak Ranch Reserve). Seedlings grew in the greenhouse for about eight weeks until all plants had their first two true leaves emerge and lengthen. Ideally, we would have placed seeds directly into the field, but to maximize biosafety, we used seedling transplants that could be tracked with 100% certainty.

We measured the longest leaf for each plant and then transplanted them into the ground in late February 2021 at Blue Oak Ranch Reserve. During the first month of growth, we replaced any D. graveolens that died. We conducted weekly phenology surveys to assess D. graveolens plant health, and at the first sign of buds, we measured plant height and harvested the aboveground biomass by cutting at the root crown and drying in a 60ºC oven for 3 days before weighing.

#Data Analysis
Statistical analyses were performed in R version 4.2.1 (R Core Team 2022) using linear mixed-effects models with the lme4 (Bates et al. 2015), lmerTest (Kuznetsova et al. 2017), and DHARMa packages (Hartig 2022), generalized linear mixed models with the glmmTMB package (Brooks et al. 2017), and mixed effects cox models with the coxme (Therneau 2022a) and survival (Therneau 2022b) packages.





  
##Generalized linear mixed models
under construction


#Libraries
```{r}
#install.packages("coxme")
#install.packages("survival")
#install.packages("ggplot2")
#install.packages("ggfortify")
#install.packages("car")
#install.packages("multcomp")
#install.packages("lme4")
#install.packages("lmerTest")
#install.packages("DHARMa")
#install.packages("dplyr")
#install.packages("emmeans")
#install.packages('TMB', type = 'source')
#install.packages("glmmTMB")
#install.packages("MASS")
#install.packages("emmeans")
#install.packages("AICcmodavg")
library(coxme)
library(survival)
library(ggplot2)
library(ggfortify)
library(car)
library(multcomp)
library(lme4)
library(lmerTest)
library(DHARMa)
library(dplyr)
library(emmeans)
library(TMB)
library(glmmTMB)
library(MASS)
library(emmeans)
library(AICcmodavg)
library(tidyverse)
```

#Load Data
This dataframe has one row per plant (800 observations). Data are for survivorship curves (3 censor options), the number of days the plant stayed alive (NumDaysAlive) and aboveground biomass. Censors with a 1 denote reaching the event (CensorAll = died, CensorBiomass = survived to collect biomass, CensorReproduction = survived to reproduce) and a 0 denoting when a seed didn't germinate by the last census date (Census = 11/15/21). CensorReproduction will be most useful in understanding the amount of biomass produced by an individual when buds appear.
```{r}
mydata-read.csv("/Users/Miranda/Documents/Education/UC Santa Cruz/Dittrichia/Dittrichia_Analysis/data/field_relative_fitness/field_relative_fitness_2021.csv",stringsAsFactors=T)
str(mydata) #Check that each column has the right class (factor, integer, numeric, etc.)
mydata$Site<-as.character(mydata$Site)
mydata$Treatment<-factor(mydata$Treatment,
                         levels=c("Grassland","Biomass Removal","Hemizonia","Raking","Raking & Clipping")) #Changing the contrast order so that everything is compared to Grassland (control)
```

#Early Growth
This code uses Growth data with Habitat (roadside and vegetated) and Treatment in a glmm model. Anova and Tukey tests are used on the successful Model4 with the creation of a box plot as a finished product.

Note: 10 blocks, 5 treatments, 16 populations (CHE-O), 8 sites (random: population pairs; CHE), 2 habitat (fixed effect: roadside and vegetated; R or O)

Also Note: Growth data is a measure of growth of longest leaf. Each plant was measured upon transplanting into the field in March and then again in June 2021. This plant has a juvenile stage of a basal rosette, and then it bolts and produced smaller cauline leaves. The goal was to capture early growth data for this plant before bolting occurs so that the measurements capture the basal rosette stage, however in some cases the plants bolted earlier than expected and the resulting measurement was smaller than previous. In these cases we determined that the data should be removed from the dataset as the negative number (or changing it to a zero) does not reflect the biological importance of the measurement.

Here we need to filter the data to remove Growth>0 and to convert plant measurement dates to date format
```{r}
growth_mydata<-mydata%>%filter(Growth>0) #Here I am only looking at the Growth data that is greater than 0 (see Also Note above)
growth_mydata$PlantDate<-as.Date(growth_mydata$PlantDate,"%m/%d/%y") 
growth_mydata$Num.Days.Growth<-as.Date("2021-05-22")-growth_mydata$PlantDate
growth_mydata$Num.Days.Growth<-as.numeric(growth_mydata$Num.Days.Growth)
growth_mydata$Growth.Rate<-growth_mydata$Growth/growth_mydata$Num.Days.Growth #First calculate number of days of growth to get the Rate (Growth/Num.Days.Growing). Then fit data to Beta distribution
```

##Histograms
###Original data
When we plot the original data as a histogram, we find that it is skewed. We also want to test for normality by using the Shapiro-Wilk normality test.

The Shapiro-Wilk normality test is used to assess whether a dataset follows a normal distribution. In the provided output, the test statistic (W) is 0.90445 and the p-value is less than 2.2e-16.

In statistical hypothesis testing, a common significance level (alpha) is set at 0.05. If the p-value is less than alpha (p-value < alpha), it suggests strong evidence against the null hypothesis. In this case, the p-value is extremely small (less than 2.2e-16), indicating very strong evidence against the null hypothesis.

Therefore, based on the given output, it can be concluded that the dataset did not pass the Shapiro-Wilk normality test. The data is unlikely to follow a normal distribution. Let's consider log transforming the data.
```{r eval=FALSE, include=FALSE}
hist(growth_mydata$Growth.Rate,
     col='steelblue',main='Original') 
shapiro.test(growth_mydata$Growth.Rate)
```

###Log transform data
(https://www.statology.org/transform-data-in-r/)
When we log transform the data and plot using a histogram, the data looks better than the original distribution, but we need to still test for normalicy with a Shapiro-Wilk normality test. Here we find that the data are not normally distributed because the p-value = 1.503e-15.
```{r eval=FALSE, include=FALSE}
log_growth_mydata<-log10(growth_mydata$Growth.Rate)
hist(log_growth_mydata,
     col='steelblue',main='Log Transformed') #Log transformed data, this looks better than the original distribution
shapiro.test(log_growth_mydata) #Data does not improve with log transformation
```

###Square root transform data
This improved the distribution, but it failed the Shapiro-Wilk normality test (p-value = 7.695e-05). Let's try it anyways with Model 1.
```{r eval=FALSE, include=FALSE}
sqrt_growth_mydata<-sqrt(growth_mydata$Growth.Rate)
hist(sqrt_growth_mydata,
     col='steelblue',main='Log Transformed') #Square root transformed data looks better than the original distribution
shapiro.test(sqrt_growth_mydata)
```

##Models
Linear Mixed Effects models are used for regression analyses involving dependent data. For a tutorial: https://doi.org/10.1177/2515245920960351

fullmodel<-lmer(log(Biomass)~   #Response variable: biomass
          Habitat*Treatment+    #Fixed effects and their interactions(*)
          (1|Site)+(1|Block),   #Random effect with random intercept only
          data=mydata)          #Dataframe
          
The linear mixed-effects model is a statistical model that incorporates both fixed effects (predictors) and random effects (grouping factors) to analyze data with hierarchical or clustered structures. It is particularly useful when dealing with correlated data, where observations within the same group may be more similar to each other than to observations from other groups.

The lmer() function allows you to specify a linear model formula, similar to the lm() function, with both fixed and random effects. It estimates the fixed effects coefficients and random effects variance components, allowing for the modeling of the relationship between the response variable and the predictors while accounting for the hierarchical structure of the data.

The random effects in lmer() capture the variability between different groups or subjects and can account for correlations among observations within the same group. This makes it suitable for analyzing longitudinal or repeated measures data, clustered data, and nested data, among others.
          
###Model 1 - lmer: Modeling Habitat and Treatment with Site and Block as random effects
For this first model, we use the sqrt data, but it fails the singularity test*. Further checking shows that Site as a random effect does not explain any of the variance in the model, therefore let's try Site as a fixed effect to see if it adds to the model.

*The message "boundary (singular) fit: see help('isSingular')" typically indicates an issue with the model fit. It suggests that the model may be encountering singularity or perfect separation.

Singularity refers to a situation where one or more predictor variables in the model are perfectly correlated or linearly dependent, leading to an unstable or unreliable estimation of the model parameters. Perfect separation occurs when a predictor variable or combination of variables can perfectly predict the outcome variable, resulting in infinite parameter estimates.

To address this problem, you may need to review your model specification, check for collinearity among predictor variables, consider removing or transforming variables, or use alternative modeling techniques suited for dealing with singularity or separation issues.
```{r eval=FALSE, include=FALSE}
growth_fullmodel1<-lmer(sqrt(Growth.Rate)~
                          Habitat*Treatment+
                          (1|Site)+
                          (1|Block),
                        data=growth_mydata)
isSingular(growth_fullmodel1,
           tol=1e-4) #=True
summary(growth_fullmodel1) #Variance explained by Site = 0.000
anova(growth_fullmodel1)
```

###Model 2 - lmer: Modeling Site as a fixed effect and Block as a random effect
Here we go back to using the log(Growth.Rate) data and find that Site as a fixed effect accounts for 3% of the variance. *** I'm not sure why I switched back to log, when my notes clearly show that log and sqrt both fail the Shapiro-Wilk test of normality
```{r eval=FALSE, include=FALSE}
growth_fullmodel2<-lmer(log(Growth.Rate)~
                          Habitat*Treatment+
                          Site+
                          (1|Block),
                        data=growth_mydata)
summary(growth_fullmodel2) #As a fixed effect, one of the Sites (OAP) is significant.
anova(growth_fullmodel2) #Site as a fixed effect accounts for 3% of the variance
```

Now we'll look at the QQ plots and the residuals using DHARMa. The QQ plots are not looking good for this model because we don't want any text or lines to be red. So, the next step is to try fitting to a negative binomial distribution. So, we'll move on to Model 3.
```{r eval=FALSE, include=FALSE}
qqnorm(resid(growth_fullmodel2)) #qqplot
qqline(resid(growth_fullmodel2)) #add the line
testDispersion(growth_fullmodel2) #red line should be in the middle of the distribution
myDHARMagraph2<-simulateResiduals(growth_fullmodel2) #making a graph using DHARMa package, also testing for heteroscedasticity, https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html#heteroscedasticity
plot(myDHARMagraph2) #plotting graph. At this point, you don't want any text or lines to be red. The QQ plots are not looking good for this model. Let's try fitting to a negative binomial distribution.
```

###Model 3 - glmer.nb: Negative Binomial
Here I tried fitting my data to a negative binomial distribution, but it throws an error that I haven't resolved. So this model isn't working well either. Let's try building a model and fitting it to a Beta distribution, first without a link function. Check it with DHARMa, and if it doesn't look good, then fit it to a Beta distribution with a logit link function.

fullmodel<-glmer.nb(Growth.Rate)~   #Response variable: growth rate
          Habitat*Treatment+        #Fixed effects and their interactions(*)
          (1|Site)+(1|Block),       #Random effect with random intercept only
          data=mydata)              #Dataframe

Generalized linear mixed-effects model (GLMM) with a negative binomial distribution is part of the lme4 package, which is commonly used for fitting various types of mixed-effects models.

The negative binomial distribution is often used to model count data that exhibit overdispersion, meaning the variance is greater than the mean. This distribution is useful when the assumption of a Poisson distribution (equal mean and variance) is violated.

The glmer.nb() function specifically fits a GLMM with a negative binomial distribution by taking into account both fixed effects (predictors) and random effects (grouping factors). It allows for the modeling of correlated data where observations within the same group may be more similar to each other than to observations from other groups.

```{r eval=FALSE, include=FALSE}
growth_fullmodel3<-glmer.nb(Growth.Rate~
                              Habitat*Treatment+
                              (1|Site)+
                              (1|Block),
                            data=growth_mydata) #this barfs right now, not sure why
isSingular(growth_fullmodel3,
           tol=1e-4)
#summary(growth_fullmodel3)
#qqnorm(resid(growth_fullmodel3)) #qqplot
#qqline(resid(growth_fullmodel3)) #add the line
#testDispersion(growth_fullmodel3) #red line should be in the middle of the distribution
#myDHARMagraph3<-simulateResiduals(growth_fullmodel3) #making a graph using DHARMa package, also testing for heteroscedasticity, https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html#heteroscedasticity
#plot(myDHARMagraph3) #plotting graph.The QQ plots are not looking good for this model.
```

###Model 4 - glmm: Beta Distribution
Now I'm using glmm because I'm fitting to other distributions (beta). The DHARMa package looks good after verifying the outliers make sense. This is now the Best Model for this data.

fullmodel<-glmmTMB(Growth.Rate~    #Response variable: growth rate
          Habitat*Treatment+       #Fixed effects and their interactions (*)
          (1|Site)+(1|Block),      #Random effect with random intercept only
          family=beta_family(),    #Specify the family as beta_family (for beta regression)
          data=growth_mydata)      #Dataframe

Generalized linear mixed-effects models (GLMMs) using a template model builder (TMB) framework. GLMMs are a type of statistical model that extends generalized linear models (GLMs) to account for correlated or clustered data, hierarchical structures, and random effects. They are suitable for analyzing data with non-normal response variables or data that violate the assumptions of traditional linear models.

The glmmTMB() function allows you to specify a GLMM using a formula syntax similar to GLMs. It supports various families of distributions (e.g., Gaussian, binomial, Poisson) and link functions, allowing for the analysis of different types of response variables. Additionally, it allows for the inclusion of both fixed effects (predictors) and random effects (grouping factors) to capture the variability within and between groups.

The TMB framework implemented in glmmTMB utilizes efficient algorithms for model fitting and parameter estimation. It offers computational advantages, making it particularly useful for large datasets or models with complex structures. The package also provides functionality for handling zero-inflation, overdispersion, and handling non-Gaussian response distributions.

```{r}
growth_fullmodel4<-glmmTMB(Growth.Rate~
                             Habitat*Treatment+
                             (1|Site)+
                             (1|Block),
                           family=beta_family(),
                           data=growth_mydata) 
summary(growth_fullmodel4)
qqnorm(resid(growth_fullmodel4)) #qqplot
qqline(resid(growth_fullmodel4)) #add the line
testDispersion(growth_fullmodel4) #red line should be in the middle of the distribution
myDHARMagraph4<-simulateResiduals(growth_fullmodel4) #making a graph using DHARMa package, also testing for heteroscedasticity, https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html#heteroscedasticity
plot(myDHARMagraph4) #plotting graph. Looks good, but check the outliers to make sure they are real.

growth_outlier_boxplot1<-ggplot(growth_mydata)+
  geom_boxplot(aes(x=Habitat,
                   y=Growth.Rate),
               size=0.5)+
  theme_classic()+
  ylim(0,1)+
  scale_fill_manual(values=c("forestgreen","gray85"))+
  labs(y="Early Growth Rate",
       fill="Habitat",
       x="Habitat")
growth_outlier_boxplot1

max(growth_mydata$Growth.Rate[growth_mydata$Habitat=="Vegetated"])
max(growth_mydata$Growth.Rate[growth_mydata$Habitat=="Roadside"]) #Yes, these outliers make sense, so I don't need to worry about the red stars in the DHARMa plot.
```

##Best Model: Early Growth
Family: beta (logit)
Formula: Growth.Rate ~ Habitat * Treatment + (1 | Site) + (1 | Block)
Data: growth_mydata

In the field experiment, source habitats (roadside and vegetated sites) did not differ for D. graveolens growth rate (Z = 0.73, P = 0.47, Figure 4A).  Although treatment had a strong effect on growth rate (Z = 10.66, P < 0.0001), there was no interaction between treatment and source habitat (Z = -0.172, P = 0.863).
```{r}
growth_fullmodel4<-glmmTMB(Growth.Rate~
                             Habitat*Treatment+
                             (1|Site)+
                             (1|Block),
                           family=beta_family(),
                           data=growth_mydata)
summary(growth_fullmodel4)
```
###Post-Hoc Test
Remove non-significant interaction terms before running the Tukey.
```{r}
#?emmeans,emmeans(model,pairwise~treatment)
growth_fullmodel4.1<-glmmTMB(Growth.Rate~
                               Treatment+
                               (1|Site)+
                               (1|Block),
                             family=beta_family(),
                             data=growth_mydata)
summary(growth_fullmodel4.1)
emmeans(growth_fullmodel4.1,
        pairwise~Treatment)
```

#Survival Analysis
This code uses NumDaysAlive data with Habitat (roadside and vegetated) and Treatment in a Cox proportional hazards model to assess survival. Information here: https://cran.r-project.org/web/packages/coxme/vignettes/coxme.pdf. "Surv" creates a survival object to combine the days column (NumDaysAlive) and the reproductive censor column (ReproductionCensor) to be used as a response variable in a model formula. The model follows the same syntax as linear models (lm, lmer, etc). Fixed effects: "var1 * var2" will give you the interaction term and the individual variables: so "var1 + var2 + var1:var2". To add random effects, type "+ (1|random effect)". 

##Histograms
```{r}
#Number of Days Alive - All data
hist(mydata$NumDaysAlive,
     col='steelblue',
     main='Original') 

#Number of Days Alive - By Habitat
ggplot(mydata,
       aes(x=NumDaysAlive))+
  geom_histogram()+
  facet_wrap(vars(Habitat)) #Here we see that both Habitats have the same bi-modal distribution.

#Number of Days Alive - By Treatment
ggplot(mydata,
       aes(x=NumDaysAlive))+
  geom_histogram()+
  facet_wrap(vars(Treatment)) #Here we see that 4 of the treatments have the same bi-modal distribution and Biomass Removal is right skewed.
```

##Models
Cox proportional hazard models
Here we will use 'coxme' which allows you to conduct mixed effects Cox proportional hazards models. Information here: https://cran.r-project.org/web/packages/coxme/vignettes/coxme.pdf. We conducted a germination experiment using Dittrichia graveolens seeds on filter paper. "Surv" creates a survival object to combine the days column (NumDaysAlive) and the censor column (CensorReproduction) to be used as a response variable in a model formula. The model follows the same syntax as linear models (lm, lmer, etc). Fixed effects: "var1 * var2" will give you the interaction term and the individual variables: so "var1 + var2 + var1:var2". To add random effects, type "+ (1|random effect)".

Assumptions for cox models: https://www.theanalysisfactor.com/assumptions-cox-regression/

###Simple Model
Start by making a simple model with no random effects. This will be compared to the full model with random effects.
```{r}
cox_simplemodel1<-coxph(Surv(NumDaysAlive,
                             CensorReproduction)~ #Only those that survived to bud are included
                          Habitat*Treatment,
                        data=mydata)
summary(cox_simplemodel1)
print(cox_simplemodel1)
predict(cox_simplemodel1)
hist(predict(cox_simplemodel1))
```

###Full Model 1
Now make a full model using random effects
```{r}
cox_fullmodel1<-coxme(Surv(NumDaysAlive,CensorReproduction)~
                        Habitat*Treatment+
                        (1|Site)+
                        (1|Block),
                      data=mydata)
summary(cox_fullmodel1)
print(cox_fullmodel1)
predict(cox_fullmodel1)
hist(predict(cox_fullmodel1))
```

Now we can compare the models to see which model is best, in this case it is fullmodel1.
```{r}
anova(cox_simplemodel1,cox_fullmodel1) 
#See example: https://www.rdocumentation.org/packages/coxme/versions/2.2-16/topics/coxme
AIC(cox_simplemodel1,cox_fullmodel1) #But comparing the AIC scores is easiest. Keep the lower AIC score because that is considered the better model. Here it is the fullmodel1.
```
###Full Model 2
Now, let's make a model with no interaction term and then we'll compare that to the fullmodel1 which was deemed the best above.
```{r}
cox_fullmodel2<-coxme(Surv(NumDaysAlive,CensorReproduction)~
                        Habitat+
                        Treatment+
                        (1|Site)+
                        (1|Block),
                      data=mydata) 
summary(cox_fullmodel2)
```

Let's compare the the first two models to test for the significance of the term that is removed (using LR). Here we see that fullmodel2 has a lower AIC score so now it is the best model.
```{r}
anova(cox_fullmodel1,cox_fullmodel2) #Not significant
AIC(cox_fullmodel1,cox_fullmodel2) #Interaction term is not significant and the second model has a lower AIC score. So we can drop the interaction term and keep fullmodel2.
```

###Full Model 3
So, now let's add in population nested under site as a random effect
```{r}
cox_fullmodel3<-coxme(Surv(NumDaysAlive,CensorReproduction)~
                        Habitat+
                        Treatment+
                        (1|Site/Population)+
                        (1|Block),
                      data=mydata) 
summary(cox_fullmodel3)
```

Now we can compare the models to see which model is best and we see that fullmodel3 has a lower AIC score.
```{r}
anova(cox_fullmodel2,cox_fullmodel3) #Significant
AIC(cox_fullmodel2,cox_fullmodel3) #Looks like fullmodel3 is the better model because of the lower AIC score
```

##Best Model: Survival
```{r}
cox_fullmodel3<-coxme(Surv(NumDaysAlive,CensorReproduction)~
                        Habitat+
                        Treatment+
                        (1|Site/Population)+
                        (1|Block),
                      data=mydata) 
summary(cox_fullmodel3)
```

###Risk Assessment
These values are found in the model summary, but if you want to pull them out, here is how you interpret them.
1 = no effect, <1 = decreased risk of death, >1 = increased risk of death.
```{r}
exp(coef(cox_fullmodel3)) #This should be interpreted that Biomass Removal is almost 5% more likely to survive to reproduction compared to Control, and Raking + Clipping is about 2.5% more likely to survive to reproduction compared to Control.
exp(ranef(cox_fullmodel3)$Block)
exp(ranef(cox_fullmodel3)$Site) # Pretty even among all the sites
Anova(cox_fullmodel3)
```

Looks like roadside and vegetated populations are the same, so let's combine them together in a model (aka, removing the Habitat term)
```{r}
cox_fullmodel4<-coxme(Surv(NumDaysAlive,CensorReproduction)~
                        Treatment+
                        (1|Site/Population)+
                        (1|Block),
                      data=mydata) 
summary(cox_fullmodel4)
summary(cox_fullmodel3)
anova(cox_fullmodel3,cox_fullmodel4) #Significant
AIC(cox_fullmodel3,cox_fullmodel4) #Looks like fullmodel4 is the better model because the AIC score is within 2 points of each other, therefore the models are assessed the same and you should take the simpler model. But this is for another manuscript :-)
```

#Reproductive Biomass
This code uses Biomass data with Habitat (roadside and vegetated) and Treatment in a lmer model. ANOVA and Tukey are used on the successful Model 3 with the creation of a box plot as a finished product.

Note: 10 blocks, 5 treatments, 16 populations (CHE-O), 8 sites (random: population pairs; CHE), 2 habitat (fixed effect: roadside and vegetated; R or O)

Also Note: Biomass data is a measure of aboveground biomass of all individuals harvested from the site. In some cases this was before they budded, but we harvested the wilted plant in case that information was needed in the future. Here we only want to look at biomass (well, log(Biomass)) for reproductive individuals.

Here we need to subset the data to only look at Biomass when CensorReproduction = 1, so that all Biomass data is for reproductive individuals only.
```{r}
reproduction_mydata<-subset(mydata,CensorReproduction%in%c('1')) #Here I am only looking at the Biomass data where the CensorReproduction = 1
```

##Histograms
###Original data
The original data is skewed and fails the Shapiro-Wilk normality test, so we should consider a log transformation.
```{r eval=FALSE, include=FALSE}
hist(reproduction_mydata$Biomass,col='steelblue',main='Original') #Original data is skewed, let's test for normality and consider log transforming the data
shapiro.test(reproduction_mydata$Biomass) #fails the Shapiro-Wilk normality test
```

###Log transform data 
(https://www.statology.org/transform-data-in-r/)
Log transformed data has a better distribution than the original data (although it still fails the Shapiro-Wilk normality test) so we will use the log transformed data with our models.

```{r}
log_reproduction_mydata<-log10(reproduction_mydata$Biomass)
hist(log_reproduction_mydata,col='steelblue',main='Log Transformed') #Log transformed data, this looks better than the original distribution
shapiro.test(log_reproduction_mydata) #fails this test
```

##Models

###Model 1 - lmer: Modeling Habitat and Treatment with Site and Block as random effects
Site as a random effect does not explain any of the variance in the model, therefore let's try Site as a fixed effect to demonstrate that it doesn't add to the model.
```{r}
fullmodel1<-lmer(log(Biomass)~
                   Habitat*Treatment+
                   (1|Site)+
                   (1|Block),
                 data=reproduction_mydata)
isSingular(fullmodel1,tol=1e-4)
summary(fullmodel1) #Variance explained by Site = 0.000
anova(fullmodel1)
predict(fullmodel1)
hist(predict(fullmodel1,type="response"))
```

###Model 2 - lmer: Modeling Site as a fixed effect
Site as a fixed effect is not significant, therefore it should not be used as a fixed effect in addition to it not being used as a random effect.
```{r}
fullmodel2<-lmer(log(Biomass)~
                   Habitat*Treatment+
                   Site+
                   (1|Block),
                 data=reproduction_mydata)
summary(fullmodel2)
anova(fullmodel2)
```

###Model 3 - lmer: Site is removed from this model because it explains very little of the variance
```{r}
fullmodel3<-lmer(log(Biomass)~
                   Habitat*Treatment+
                   (1|Block),
                 data=reproduction_mydata)
summary(fullmodel3)
anova(fullmodel3)
```

Now we'll look at the QQ plots and the residuals using DHARMa
```{r}
qqnorm(resid(fullmodel3)) #qqplot
qqline(resid(fullmodel3)) #add the line
testDispersion(fullmodel3) #red line should be in the middle of the distribution
myDHARMagraph3<-simulateResiduals(fullmodel3) #making a graph using DHARMa package, also testing for heteroscedasticity, https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html#heteroscedasticity
plot(myDHARMagraph3) #plotting graph. At this point, you don't want any text or lines to be red.
```

When we compare the model summaries and Anova of Models 2 & 3, we see that the removal of Site doesn't impact the model. Therefore the more simple Model 3 is a better choice. But let's keep checking...

Compare AIC Scores
Now I need to compare the AIC scores for all the models to tell me which is the better model (but it will not say which fits my data better, that is why I did all the DHARMa stuff)

Using aictab to make the comparison of models and table 
```{r}
models<-list(fullmodel1,fullmodel2,fullmodel3)
mod.names<-c('Site.Random','Site.Fixed','No.Site')
aictab(cand.set=models,modnames=mod.names)
#The lowest AICc score is listed first and indicates the best fitting model, here, Model 3 (No.Site) where Site is not included. The cut-off for comparing Models is 2 units. The difference between Model 1 (Site.Random) and Model 3 (No.Site) is 2.12. So Model 3 (No.Site) is marginally better than Model 1 (Site.Random).
```

###Other Models Attempted
Other ideas that were explored to keep Site but resulted in singular error and Site explaining 0.000 of the variance: 
fullmodel4<-lmer(log(Biomass)~ Treatment + (1|Site) + (1|Block), data = mydata)
summary(fullmodel4)
fullmodel5<-lmer(log(Biomass)~ Treatment + Habitat + (1|Site) + (1|Block), data = mydata)
summary(fullmodel5)

###Best Model: Biomass
```{r}
fullmodel3<-lmer(log(Biomass)~
                   Habitat*Treatment+
                   (1|Block),
                 data=reproduction_mydata)
summary(fullmodel3)
```

####Post-Hoc Test
You should remove non-significant interaction terms before running a post-hoc test. It is difficult to judge the main effects (Habitat and Site) when you also have the interaction term present, so when it is not significant, fit a new model without it.
```{r}
fullmodel3.1<-lmer(log(Biomass)~
                     Treatment+
                     (1|Block),
                   data=reproduction_mydata)
summary(fullmodel3.1)
anova(fullmodel3.1)
qqnorm(resid(fullmodel3.1)) #qqplot
qqline(resid(fullmodel3.1)) #add the line
testDispersion(fullmodel3.1) #red line should be in the middle of the distribution
myDHARMagraph3.1<-simulateResiduals(fullmodel3.1) #testing for heteroscedasticity
plot(myDHARMagraph3.1) #plotting graph
```

```{r}
fullmodel3.2<-lmer(log(Biomass)~
                     Treatment+
                     (1|Block)+
                     (1|Population),
                   data=mydata)
summary(fullmodel3.2)
anova(fullmodel3.2)
qqnorm(resid(fullmodel3.2)) #qqplot
qqline(resid(fullmodel3.2)) #add the line
testDispersion(fullmodel3.2) #red line should be in the middle of the distribution
myDHARMagraph3.2<-simulateResiduals(fullmodel3.2) #testing for heteroscedasticity
plot(myDHARMagraph3.2) #plotting graph
```

Let's use emmeans for our Best Model (fullmodel3) and the two off-shoots.
```{r}
#?emmeans, emmeans(model, pairwise ~ treatment)
emmeans(fullmodel3,pairwise~Treatment)
emmeans(fullmodel3.1,pairwise~Treatment)
emmeans(fullmodel3.2,pairwise~Treatment)
#These models result in similar Tukey outcomes
```


#ggplot - interaction plots
Treatment (Biomass Removal and Grassland) on the x-axis, using Roadside and Vegetated data

##Growth Rate x Treatment: Growth
```{r}
pd<-position_dodge(0)

growth_mydata2<-growth_mydata%>% 
  replace_na(list(Growth.Rate=0))%>%
  filter(CensorReproduction==1)%>%
  filter(Treatment=="Biomass Removal"|Treatment=="Grassland")%>%
  group_by(Treatment,Habitat)%>%
  summarise(Mean=mean(Growth.Rate),
            SD=sd(Growth.Rate),
            N=length(Growth.Rate))%>%
  mutate(SE=SD/sqrt(N))

field_int_rate_grow_gg<-ggplot(growth_mydata2,
                        aes(x=factor(Treatment,levels=c("Biomass Removal","Grassland")),
                            y=Mean,
                            shape=Habitat,
                            group=Habitat,
                            color=Habitat))+
  coord_cartesian(ylim = c(0,0.5))+
  xlab("")+
  theme_bw(base_size=16)+
  theme(panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        axis.line=element_line(colour="black"),
        axis.title.y=ggtext::element_markdown(),
        legend.position = c(0.8, 0.8))+
 # ggtitle("Growth of Dittrichia graveolens that budded")+
  ylab("*Dittrichia graveolens* \n (Change in Leaf Length/Days)")+
  geom_line(size=1,
            position=pd)+
  geom_point(size=5,
             position=pd)+
  scale_color_manual(values=c("gray85","forestgreen"))+
  geom_errorbar(width=0.15,
                position=pd,
                aes(ymin=(Mean-SE),
                    ymax=(Mean+SE)))+
  labs(y="Change in Leaf Length per Day",
       color="Source Habitat",
       shape="Source Habitat")
field_int_rate_grow_gg

#ggsave(plot=field_int_rate_grow_gg,file="/Users/Miranda/Documents/Education/UC Santa Cruz/Dittrichia/Dittrichia_Analysis/figures/field_int_rate_grow_gg.png",width=25,height=15,units="cm",dpi=800)
```

##Survival to bud X Treatment
```{r}
pd<-position_dodge(0)

surv_mydata1<-mydata%>%
  filter(Treatment=="Biomass Removal"|Treatment=="Grassland")%>%
  group_by(Treatment,Habitat)%>%
  summarise(Mean=mean(PropBudSite),
            SD=sd(PropBudSite),
            N=length(PropBudSite))%>%
  mutate(SE=SD/sqrt(N))

field_int_surv_all_gg<-ggplot(data=surv_mydata1,
                        aes(x=factor(Treatment,levels=c("Biomass Removal","Grassland")),
                            y=Mean,
                            shape=Habitat,
                            group=Habitat,
                            color=Habitat))+
  ylab("Proportion Survival to Bud")+
  coord_cartesian(ylim=c(0,1))+
  xlab("")+
  theme_classic(base_size=16)+
  theme(panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        axis.line=element_line(colour="black"),
        legend.position = c(0.8, 0.8))+
  scale_y_continuous(name="Proportion Survival to Bud",
                     limits=c(0,1),
                     breaks=c(0.0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0))+
  geom_line(size=1,
            position=pd)+
  geom_point(size=5,
             position=pd)+
  scale_color_manual(values=c("gray85","forestgreen"))+
  geom_errorbar(width=0.15,
                position=pd,
                aes(ymin=(Mean-SE),
                    ymax=(Mean+SE)))+
  labs(y="Proportion Survival to Bud",
       fill="Source Habitat",
       color="Source Habitat",
       shape="Source Habitat")
field_int_surv_all_gg

#ggsave(plot=field_int_surv_all_gg,file="/Users/Miranda/Documents/Education/UC Santa Cruz/Dittrichia/Dittrichia_Analysis/figures/field_int_surv_all_gg.png",width=25,height=15,units="cm",dpi=800)
```

##Biomass x Treatment: Growth
```{r}
pd<-position_dodge(0)

growth_mydata1<-growth_mydata%>% 
  replace_na(list(Biomass=0))%>%
  filter(CensorReproduction==1)%>%
  filter(Treatment=="Biomass Removal"|Treatment=="Grassland")%>%
  group_by(Treatment,Habitat)%>%
  summarise(Mean=mean(Biomass),
            SD=sd(Biomass),
            N=length(Biomass))%>%
  mutate(SE=SD/sqrt(N))

field_int_bio_grow_gg<-ggplot(growth_mydata1,
                        aes(x=factor(Treatment,levels=c("Biomass Removal","Grassland")),
                            y=Mean,
                            shape=Habitat,
                            group=Habitat,
                            color=Habitat))+
  ylab("Biomass (g)")+
  coord_cartesian(ylim = c(0,10))+
  xlab("")+
  #ggtitle("Biomass of Dittrichia graveolens that budded")+
  theme_bw(base_size=16)+
  theme(panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        axis.line=element_line(colour="black"),
        axis.title.y=ggtext::element_markdown(),
        legend.position = c(0.8, 0.8))+
  geom_line(size=1,
            position=pd)+
  geom_point(size=5,
             position=pd)+
  scale_color_manual(values=c("gray85","forestgreen"))+
  geom_errorbar(width=0.15,
                position=pd,
                aes(ymin=(Mean-SE),
                    ymax=(Mean+SE)))+
  labs(y="*Dittrichia graveolens* Biomass (g)",
       color="Source Habitat",
       shape="Source Habitat")
field_int_bio_grow_gg

#ggsave(plot=field_int_bio_grow_gg,file="/Users/Miranda/Documents/Education/UC Santa Cruz/Dittrichia/Dittrichia_Analysis/figures/field_int_bio_grow_gg.png",width=25,height=15,units="cm",dpi=800)
```

##Figure panel (3 plots)
```{r}
library(patchwork)

multi_panel_gg<-(field_int_rate_grow_gg+theme(legend.position = c(0.9,0.9)))/
  (field_int_surv_all_gg+theme(legend.position="none"))/
  (field_int_bio_grow_gg+theme(legend.position="none"))+
  plot_annotation(tag_levels=c("A"),
                  tag_suffix=") ")&
  theme(text=element_text(size=12))
multi_panel_gg

ggsave(plot=multi_panel_gg,file="/Users/Miranda/Documents/Education/UC Santa Cruz/Dittrichia/Dittrichia_Analysis/figures/multi_panel_gg.png",width=15,height=20,units="cm",dpi=800)
```