---
title: "seedbanks"
output: html_document
author: Emma Gloudeman, Miranda Melen
date: May 2023
---

#Meta Data
This is year 1 and 2 data from the greenhouse raised bed seedbank experiment.

#Packages
```{r}
library(ggplot2)
library(magrittr)
library(tidyverse)
```

#Load Data
```{r}
gh_bank_all<-read.csv("/Users/Miranda/Documents/Education/UC Santa Cruz/Dittrichia/Dittrichia_Analysis/data/seedbanks/greenhouse_seedbank_data.csv",stringsAsFactors=T)

rain_data<-na.omit(read.csv("/Users/Miranda/Documents/Education/UC Santa Cruz/Dittrichia/Dittrichia_Analysis/data/seedbanks/younger_lagoon_rainfall-hourly.csv"))
```

##Filter = 200 seeds
Not every pot received 200 seeds, so here we will filter to only include those that did, to simplify for now.
```{r}
gh_bank<-gh_bank_all[gh_bank_all$SeedCount==200,] #filters a data frame gh_bank_all and creates a new data frame gh_bank that contains only the rows where the value in the SeedCount column is equal to 200.
#view(gh_bank)
```

##Make a new dataframe
Here we will calculate the mean germination for each month within the gh_bank data frame and assign them to the prop_means vector
```{r}
prop_means=c(mean(gh_bank$Sep1_prop), mean(gh_bank$Oct1_prop), mean(gh_bank$Nov1_prop), 
    mean(gh_bank$Dec1_prop),mean(gh_bank$Jan1_prop),mean(gh_bank$Feb1_prop),
    mean(gh_bank$Mar1_prop),mean(gh_bank$Apr1_prop),mean(gh_bank$May1_prop),
    mean(gh_bank$Jun1_prop),mean(gh_bank$Jul1_prop),mean(gh_bank$Aug1_prop),
    mean(gh_bank$Sep2_prop),mean(gh_bank$Oct2_prop),mean(gh_bank$Nov2_prop), 
    mean(gh_bank$Dec2_prop),mean(gh_bank$Jan2_prop),mean(gh_bank$Feb2_prop),
    mean(gh_bank$Mar2_prop))

#Make months
months=c('Sep21','Oct21','Nov21','Dec21','Jan22','Feb22','Mar22','Apr22',
         'May22','Jun22','Jul22','Aug22',
         'Sep22','Oct22','Nov22','Dec22','Jan23','Feb23','Mar23')
```

##Summarize rain data
```{r}
rain_sum<-rain_data%>% 
  group_by(MonthNum,Year)%>%
  summarise(sum=sum(Rainfallmm))
#View(rain_sum)

rain_sum<-rain_sum[order(rain_sum$Year,rain_sum$MonthNum),]


rain<-c(rain_sum$sum[c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19)])
```

##Combine data into new datafram
Reorder the months
```{r}
gh_bank_summary<-data.frame(prop_means,months,rain)
#View(gh_bank_summary)

gh_bank_summary$months<-ordered(gh_bank_summary$months, levels=c('Sep21','Oct21','Nov21','Dec21','Jan22','Feb22','Mar22','Apr22','May22','Jun22','Jul22','Aug22',                                                        'Sep22','Oct22','Nov22','Dec22','Jan23','Feb23','Mar23'))
```

#Figure with two y-axis (only data with 200 seeds)
```{r}
#set the scale for the right y-axis to 250
scale <- 250

gh_bank_gg<-ggplot(data=gh_bank_summary,
                   aes(x=months,
                       y=prop_means))+
  geom_line(group=1,
            color="forestgreen",
            linewidth=1.2)+
  geom_point(size=3,
             color="forestgreen")+
  xlab("Months")+
  ylab("Proportion Germinated")+
  scale_y_continuous(limits=c(0,1),
                     breaks=c(0.0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0),
                     sec.axis=sec_axis(~.*scale,name="Rainfall (mm)"))+
  theme_bw(base_size=16)+
  theme(panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        axis.line=element_line(colour="black"),
        legend.position=c(0.8, 0.4))+
  geom_bar(data=gh_bank_summary,
  aes(x=months,
  y=rain/scale),
         stat="identity",
         fill="blue",
         alpha=0.5)
gh_bank_gg

ggsave(plot=gh_bank_gg,file="/Users/Miranda/Documents/Education/UC Santa Cruz/Dittrichia/Dittrichia_Analysis/figures/gh_bank_gg.png",width=25,height=15,units="cm",dpi=800)
```